---
title: "R Notebook"
output:
  html_notebook: default
  pdf_document:
    latex_engine: xelatex
always_allow_html: yes
---

```{r setup, include=FALSE}

# This chunk shows/hides the code in your final report. When echo = TRUE, the code
# is shown in the report. When echo = FALSE, the code is hidden from the final report.
# We would like to see your code, so please leave the setting as is during the course.
# This chunk will not show up in your reports, so you can safely ignore its existence.

knitr::opts_chunk$set(echo = TRUE)

```


The following is your first chunk to start with. Remember, you can add chunks using the menu
above (Insert -> R) or using the keyboard shortcut Ctrl+Alt+I. A good practice is to use
different code chunks to answer different questions. You can delete this comment if you like.

Other useful keyboard shortcuts include Alt- for the assignment operator, and Ctrl+Shift+M
for the pipe operator. You can delete these reminders if you don't want them in your report.


```{r}
#setwd("...")

library("tidyverse")
library("tidymodels")
library("plotly")
library("skimr")
```

```{r}
library("gapminder")
library("ggrepel")
```

```{r}
#-----Question 2(a)-----
dfGap <- gapminder

```

```{r}
#-----Question 3(a)-----
skim(dfGap)

```

```{r}
#-----Question 3(b)-----
dfGap %>%
  filter(year == 2007) %>%
  arrange(desc(lifeExp))

```

```{r}
#-----Question 3(c)-----
dfGap %>%
  mutate(totalGDP = pop * gdpPercap) %>% 
  filter(year == 2007) %>% 
  arrange(desc(totalGDP))

```

```{r}
#-----Question 3(c(iii))-----
dfGap %>% 
  filter(year == 2007) %>%
  select(country, gdpPercap) %>% 
  arrange(desc(gdpPercap))

```

```{r}
#-----Question 3(d)-----
dfGap %>%
  mutate(totalGDP = pop * gdpPercap) %>%
  filter(year == 2007) %>% 
  group_by(continent) %>%
  summarize(medianLifeExp = median(lifeExp), medianTotalGDP = median(totalGDP)) %>%
  ungroup() %>% 
  arrange(desc(medianLifeExp))

```

```{r}
#-----Question 4(a(i))-----
dfGap %>%
  mutate(totalGDP = pop * gdpPercap) %>% 
  filter(year == 2007) %>% 
  ggplot(aes(x = totalGDP, y = lifeExp)) + geom_point()

```

```{r}
#-----Question 4(a(ii))-----
dfGap %>%
  mutate(totalGDP = pop * gdpPercap) %>% 
  filter(year == 2007, totalGDP > 1e+12) %>% 
  ggplot(aes(x = totalGDP, y = lifeExp)) +
  geom_point()

```

```{r}
#-----Question 4(a(iii))-----
dfGap %>%
  mutate(totalGDP = pop * gdpPercap) %>% 
  filter(year == 2007, totalGDP > 1e+12) %>% 
  ggplot(aes(x = totalGDP, y = lifeExp)) +
  geom_point() + geom_label(aes(label = country))

```

```{r}
#-----Question 4(a(iv))-----
dfGap %>%
  mutate(totalGDP = pop * gdpPercap) %>% 
  filter(year == 2007, totalGDP > 1e+12) %>% 
  ggplot(aes(x = totalGDP, y = lifeExp)) +
  geom_point() + geom_label_repel(aes(label = country))

```

```{r}
#-----Question 4(a(v))-----
dfGap %>%
  mutate(totalGDP = pop * gdpPercap) %>% 
  filter(year == 2007, totalGDP > 1e+12) %>% 
  ggplot(aes(x = totalGDP, y = lifeExp)) +
  geom_point(aes(color = continent)) + geom_label_repel(aes(label = country))

```

```{r}
#-----Question 4(b)-----
dfGap %>%
  filter(year == 2007) %>% 
  ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point(aes(color = continent)) + geom_smooth()

```

```{r}
#-----Question 4(c)-----
dfGap %>% 
  ggplot(aes(x = continent, y = lifeExp)) + geom_boxplot() +
  ggtitle("Comparing life expectancy across continents")

```

```{r}
#-----Question 4(c(iii))-----
boxPlotsForAll <- 
  dfGap %>% 
  ggplot(aes(x = continent, y = lifeExp)) + geom_boxplot() +
  ggtitle("Comparing life expectancy across continents")

ggplotly(boxPlotsForAll)

```

```{r}
#-----Question 4(d(i))-----
dfGap %>%
  group_by(year, continent) %>%
  summarize(medianGDPPerCap = median(gdpPercap)) %>% 
  ggplot(aes(x = year, y = medianGDPPerCap, color = continent)) + geom_line()

```

```{r}
#-----Question 4(d(i))----- ALTERNATIVE, ONE-LINE LONGER, SOLUTION
dfGap %>%
  group_by(year, continent) %>%
  mutate(medianGDPPerCap=median(gdpPercap))%>%
  distinct(continent, .keep_all=TRUE) %>% 
  ggplot(aes(x = year, y = medianGDPPerCap, color = continent)) + geom_line()

```


```{r}
#-----Question 4(d(ii))-----
dfGap %>%
  group_by(year, continent) %>%
  summarize(meanGDPPerCap = mean(gdpPercap)) %>% 
  ggplot(aes(x = year, y = meanGDPPerCap, color = continent)) + geom_line()

```

```{r}
#-----Question 4(d(iii))-----
gdpOverTime <- 
  dfGap %>%
  group_by(year, continent) %>%
  summarize(medianGDPPerCap = median(gdpPercap)) %>% 
  ggplot(aes(x = year, y = medianGDPPerCap, color = continent)) + geom_line()

ggplotly(gdpOverTime)

```