---
title: "R Notebook"
output:
  pdf_document:
    latex_engine: xelatex
  html_notebook: default
always_allow_html: yes
---

```{r setup, include=FALSE}

# This chunk shows/hides the code in your final report. When echo = TRUE, the code
# is shown in the report. When echo = FALSE, the code is hidden from the final report.
# We would like to see your code, so please leave the setting as is during the course.
# This chunk will not show up in your reports, so you can safely ignore its existence.

knitr::opts_chunk$set(echo = TRUE)

```


The following is your first chunk to start with. Remember, you can add chunks using the menu
above (Insert -> R) or using the keyboard shortcut Ctrl+Alt+I. A good practice is to use
different code chunks to answer different questions. You can delete this comment if you like.

Other useful keyboard shortcuts include Alt- for the assignment operator, and Ctrl+Shift+M
for the pipe operator. You can delete these reminders if you don't want them in your report.


```{r}
#setwd("C:/") #Don't forget to set your working directory before you start!

library("tidyverse")
library("tidymodels")
library("plotly")
library("skimr")
library(lubridate)
```

```{r}
dfbOrg <-
 read_csv("assignment2BikeShare.csv") 
```

```{r}
skim(dfbOrg)
```

```{r}
#Q1a. i)
dfbOrg <-dfbOrg %>%
 mutate(COUNT= CASUAL+ REGISTERED)
dfbOrg
```

```{r}
#Q1a. ii)
y <- months(dfbOrg$DATE,abbr =TRUE)
dfbOrg <-dfbOrg %>%
 mutate(MONTH= y)
dfbOrg
```
```{r}
#Q1b.
dfbStd <-dfbOrg %>%
 mutate(TEMP= scale(TEMP, center = TRUE, scale = TRUE)) %>%
 mutate(ATEMP= scale(ATEMP, center = TRUE, scale = TRUE)) %>%
 mutate(HUMIDITY= scale(HUMIDITY, center = TRUE, scale = TRUE)) %>%
 mutate(WINDSPEED= scale(WINDSPEED, center = TRUE, scale = TRUE))
dfbStd
```
```{r}
#Q2.
fitAll <- lm(COUNT~., data=dfbStd)
summary(fitAll)
```
```{r}
#Q3 a.
dfbOrg <- dfbOrg %>%
 mutate(BADWEATHER = ifelse(WEATHERSIT==3| WEATHERSIT==4 , 'YES', 'NO'))
dfbOrg
```
```{r}
#Q3 b
plot <- dfbOrg %>%
ggplot(mapping= aes(x=ATEMP,y=COUNT,color= BADWEATHER))+geom_point()
plot
```
```{r}

plot2 <- 
  dfbOrg %>% 
    ggplot(aes(x=ATEMP, y=CASUAL, color=factor(BADWEATHER)))+geom_point()

plot2
```


```{r}
#Q3 c. i)
plot3 <- dfbOrg %>%
ggplot(mapping= aes(x=ATEMP,y=REGISTERED,color= factor(BADWEATHER) ))+geom_point()
plot3
```
```{r}
#Q3 c. iv)
plot4 <- dfbOrg %>%
ggplot(mapping= aes(x=ATEMP,y=COUNT ))+
  geom_point()+geom_smooth()
plot4
## `geom_smooth()` using method = 'loess' and formula 'y ~ x'
```
```{r}
#Q4
fit4 <- lm(formula = COUNT ~ MONTH + WEEKDAY + BADWEATHER + TEMP + ATEMP + HUMIDITY  ,data = dfbOrg)
summary(fit4)

```
```{r}
#Q5
plot(fit4)
```
```{r}
#Q6 a
bvc <- lm(formula = COUNT ~ BADWEATHER, data=dfbOrg)
summary(bvc)
```
```{r}
#Q6 b.
badweek <- lm(formula=COUNT~ BADWEATHER +WEEKDAY , data=dfbOrg)
summary(badweek)
```

```{r}
#Q7
set.seed(333)
dfbTrain <- dfbOrg %>%
sample_frac(0.8)

# Model 1
dfbTest <- dplyr::setdiff(dfbOrg, dfbTrain)
fitOrg <-lm(formula=COUNT~ MONTH+WEEKDAY+BADWEATHER+TEMP+ATEMP+HUMIDITY,
data=dfbTrain)
summary(fitOrg)
```
```{r}
resultsOrg <-dfbTest %>%
mutate(predictedCount = predict(fitOrg, dfbTest))
resultsOrg
```
```{r}
performance <- metric_set(rmse, mae)
performance(resultsOrg, truth= COUNT, estimate=predictedCount)
```
```{r}
set.seed(333)
dfnwTrain <- dfbOrg %>%
sample_frac(0.8)
dfnwTest <- dplyr::setdiff(dfbOrg, dfnwTrain)

#Model 2
fitNew <-lm(formula=COUNT~
MONTH+WEEKDAY+BADWEATHER+TEMP+ATEMP+HUMIDITY+WINDSPEED, data=dfbTrain)
summary(fitNew)
```
```{r}
resultsNew <-dfnwTest %>%
mutate(predCnt = predict(fitNew, dfnwTest))
resultsNew
```
```{r}
performance <- metric_set(rmse, mae)
performance(resultsNew, truth= COUNT, estimate=predCnt)

```
```{r}
#Q8 
dfbTrain2 <- dfbOrg %>%
 filter(year(DATE)==2011)
dfbTest2 <- dfbOrg %>%
 filter(year(DATE)==2012) 
```
```{r}
fitOrg2 <-
 lm(COUNT ~ MONTH+WEEKDAY+BADWEATHER+ATEMP+HUMIDITY, dfbTrain2)
summary(fitOrg2)
```
```{r}
lm(formula = COUNT ~ MONTH + WEEKDAY + BADWEATHER + ATEMP + HUMIDITY,
 data = dfbTrain2)
```
```{r}
resultsOrg2 <-
 dfbTest2 %>%
 mutate(predictedCnt = predict(fitOrg2, dfbTest2))
resultsOrg2
```

```{r}
performance <-
 metric_set(rmse, mae)
performance(resultsOrg2, truth=COUNT, estimate=predictedCnt)
```
```{r}
dfbTest3 <-
 dfbOrg %>%
 filter(year(DATE)==2012) %>%
 filter(month(DATE)>06)
dfbTrain3 <- dplyr::setdiff(dfbOrg, dfbTest3)
```

```{r}
fitOrg3 <-
 lm(COUNT ~ MONTH+WEEKDAY+BADWEATHER+ATEMP+HUMIDITY, dfbTrain3)
summary(fitOrg3)
```
```{r}
resultsOrg3 <-
 dfbTest3 %>%
 mutate(predictedCnt = predict(fitOrg3, dfbTest3))
resultsOrg3
```
```{r}
performance <-
 metric_set(rmse, mae)
performance(resultsOrg3, truth=COUNT, estimate=predictedCnt)
```

